A Lattice Boltzmann Method for immiscible multiphase flow simulations using the Level Set Method

### 2015-08-19 Vortrag/folien.tex
noch einmal Folien...

\begin{frame}
To solve the two-phase problem we need to:
\begin{itemize}
\item<1-> solve the flow equations $\rightarrow$ LBM
\item<2-> compute the motion of the interface $\rightarrow$ Level Set Method
\item<3-> couple the two algorithms $\rightarrow$ according BC's % Fuers coupling muessen ja eigentlich nur eine spezielle LBM-Randbehandlung am Interface gemacht werden, und fuer Coupling in die anderere Richtung halt die Geschwindigkeit interpoliert werden
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface capturing}
The interface between fluid phases is captured by a Level-Set Method. 
I.e. a \textit{level set function} $\varphi:= \varphi(x,t) \rightarrow \mathbb{R}$ is tracked through the fluid domain. The interface is given (implicitly) by the zero-isosurface of this function.
It holds:
$$ \frac{\partial \varphi}{\partial t} + \vec v \cdot \nabla \varphi = 0$$
where $\vec v$ is the velocity at the interface; given by NSE.

Benefit of Level-Set Method:
\begin{itemize}
\item<2-> Implicit surface description eases bubble coalescence/breakup in code
\item<3-> High density and viscosity ratios ($>10^3$) possible for simulation
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface implementation (Methods Coupling)}
Hydrodynamics are solved by LBM (Stream-Collide, D2Q9, SRT for our first test)
\begin{itemize}
\item<1-> Interface becomes a new boundary condition for LBM
  $$f_i(x,t+1) = f_{i^*}^{+}(x,t) + 6 \Delta x f_{i}^{*}c_i \cdot \tilde{v} + R_i$$
\end{itemize}
\begin{tabular}{l l}
\hspace{-.35cm} \begin{minipage}{0.65\textwidth}
\visible<1->{
\begin{itemize}
\item<2-> $\tilde{v}$ is the velocity on the interface along the direction $c_i$
\item<3-> $R_i$ ensures the jump conditions of the normal stress and corrects the error terms resulting from the bounce back treatment
\end{itemize}
}
\end{minipage}
&
\hspace{-.5cm} \begin{minipage}{0.4\textwidth}
\begin{figure}[h!]
\includegraphics[width=3cm]{skizze2.png}
\end{figure}
\end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\visible<1->{
$\tilde{v}$ is calculated by linear interpolation:
$$\tilde{v} = q v (x_2,t) + (1-q) v(x_1,t)$$
}
\visible<2->{
$R_i$ consists of several parts:
$$R_i = 6 \Delta x^2 f_{i}^{*} {\Lambda}_i : A$$
}
\visible<3->{
\begin{itemize}
\item<3-> with:
$$ {\Lambda}_i = c_i \otimes c_i - \frac13 {|c_i|}^2 \mathbb{I}$$
$$ A = -q(1-q)[S] - (q - 1/2) S^{2}$$
\item<4-> $S^{(k)}$ velocity gradient at $x_k$
\item<5-> $[S] = \lim_{\epsilon \to 0}(S(x+\epsilon \vec n) -S(x-\epsilon \vec n))$ jump of velocity gradient at the interface. Depends on normal, tangent and curvature.
\end{itemize}
}
\end{frame}

@@

\section{Results}
\begin{frame}
\frametitle{Results}
\framesubtitle{Setup}
Transient couette flow with immersed drop.
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}

@@

\begin{frame}

\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}

@@

\begin{frame}
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
\includegraphics[width=7cm]{plots/vel10.png}
\end{figure}
\end{minipage}
&
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
  \includegraphics[width=4cm]{plots/interface_t10.png}
\end{figure}
\end{minipage}
\end{tabular}
\dots more to come 
\end{frame}

>>>>

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Solvers for the two-phase problem}
\begin{itemize}
\item<1-> Colour gradient method of Gunstensen and Rothman
\begin{itemize}
\item only for small density and viscosity differences
\end{itemize}
\item<2-> Method of Shan and Chen
\begin{itemize}
\item models miscible fluids
\item immiscible flows only approximatly
\end{itemize}
\item<3-> Free surface methods
\begin{itemize}
\item only for big density and viscosity differences
\end{itemize}
\item<4-> Method of Thoemmes and Becker
\begin{itemize}
\item using Level Set Method for interface motion
\item small and big density/viscosity differences possible
\item mass loss in Level Set Method
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Ingredients for solver with Level Set Method}
To solve the two-phase problem we need to:
\begin{itemize}
\item<1-> solve the flow equations $\rightarrow$ LBM
\item<2-> compute the motion of the interface $\rightarrow$ Level Set Method
\item<3-> couple the two algorithms $\rightarrow$ according BC's % Fuers coupling muessen ja eigentlich nur eine spezielle LBM-Randbehandlung am Interface gemacht werden, und fuer Coupling in die anderere Richtung halt die Geschwindigkeit interpoliert werden
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{LBM}
\begin{itemize}
\item<1-> Stream
$$f_i(x+c_i,t+1) = f_i(x,t) - \frac{1}{\tau}(f_i - f_i^{eq}) + G_i$$
\item<2-> Collide (Bhatnagar-Gross-Krook approximation)
$$f_i^{eq}(\rho,v) = f_i^{*}(\rho + 3c_i \cdot v + \frac{9}{2}(c_i \cdot v)^2 - \frac{3}{2}v^2)$$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{LBM}
the corresponding D2Q9 weight factors
\begin{figure}[h!]
    %\flushright
\includegraphics[width=3cm]{tutorial_basics03_d2q9.png}
\end{figure}
$$f_i^* = \begin{cases} \frac49 & i = C \\ \frac19 & i = N,E,S,W \\ \frac{1}{36} &  i = NE,SE,SW,NW \end{cases}$$
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{LBM}
we can compute for
\begin{itemize}
\item Density
$$\rho(x,t) = \sum f_i(x,t)$$
\item Velocity
$$u = \sum f_i(x,t)c_i$$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface capturing}
\visible<1->{Interface between fluid phases is captured by level set method}
\visible<2->{
\begin{itemize}
\item<2-> Level set function: $\varphi:= \varphi(x,t) \rightarrow \mathbb{R}$

Interface is the zero-isosurface of this function
\item<3-> Level set equation:
$$ \frac{\partial \varphi}{\partial t} + \vec v \cdot \nabla \varphi = 0$$
where $\vec v$ is the velocity at the interface, given by NSE.
\end{itemize}
}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface capturing}
Benefits of Level-Set Method:
\begin{itemize}
\item<2-> Implicit surface description eases bubble coalescence/breakup in code
\item<3-> High density and viscosity ratios ($>10^3$) possible for simulation
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface implementation (Methods Coupling)}
Hydrodynamics are solved by LBM (Stream-Collide, D2Q9, SRT for our first test)
\begin{itemize}
\item<1-> Interface becomes a new boundary condition for LBM
  $$f_i(x,t+1) = f_{i^*}^{+}(x,t) + 6 \Delta x f_{i}^{*}c_i \cdot \tilde{v} + R_i$$
\end{itemize}
\begin{tabular}{l l}
\hspace{-.35cm} \begin{minipage}{0.65\textwidth}
\visible<1->{
\begin{itemize}
\item<2-> $\tilde{v}$ is the velocity on the interface along the direction $c_i$
\item<3-> calculated by linear interpolation:
$$\tilde{v} = q v (x_2,t) + (1-q) v(x_1,t)$$
\end{itemize}
}
\end{minipage}
&
\hspace{-.5cm} \begin{minipage}{0.4\textwidth}
\begin{figure}[h!]
\includegraphics[width=3cm]{skizze2.png}
\end{figure}
\end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface implementation (Methods Coupling)}
$R_i$ ensures the jump conditions of the normal stress and corrects the error terms resulting from the bounce back treatment.
\begin{itemize}
\item<1-> $R_i = 6 \Delta x^2 f_{i}^{*} {\Lambda}_i : A$,

with $A : B = \sum_{i,j} A_{ij}B_{ij}$.
\item<2->
$ {\Lambda}_i = c_i \otimes c_i - \frac13 {|c_i|}^2 \mathbb{I}$ ,

with $c_i \otimes c_i = c_i c_i^T$.

\item<3->{$ A = -q(1-q)[S] - (q - 1/2) S^{(k)}$
\begin{itemize}
\item<3-> $S^{(k)}$ velocity gradient at $x_k$
\item<4-> $[S]$ jump of velocity gradient at the interface. Depends on $\vec n$, $\vec t$ and $\kappa$.
\end{itemize}
}
\end{itemize}
\end{frame}

@@

\section{Results}
\begin{frame}
\frametitle{Results}
\framesubtitle{Implementation}
\begin{itemize}
\item<1-> Implementation in Matlab
\item<2-> Own LBM-solver
\item<3-> Level set method solved by Toolbox of Ian Mitchell
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Results}
\framesubtitle{Setup}
\begin{itemize}
\item<1-> Transient couette flow with immersed drop
\item<2-> Fluids have the same $\rho_i, \nu_i \rightarrow [S]$ vanishes
\item<3-> Goal: See interactions at the interface
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Results}
\framesubtitle{Simulation}
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Simulation}
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Simulation}
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
\includegraphics[width=7cm]{plots/vel10.png}
\end{figure}
\end{minipage}
&
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
  \includegraphics[width=4cm]{plots/interface_t10.png}
\end{figure}
\end{minipage}
\end{tabular}
\end{frame}
